---
title: "STA 4102 - Homework 2"
author: "Kayla Perry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 2

## Complete All of the Problems
Complete all of the problems below and enter your code within the supplied code block. If code is already included in the code block, this is for the setup of the problem. If the problem does not have a code component, write your response as a comment in the R code block.

### Problem 1
Create a vector named first_vector with the numbers from 1 to 10 using R.
```{r problem_1, error=TRUE}
#Your code here
first_vector <- c(1:10)
print(first_vector)
```

### Problem 2
What is the difference between 1 and 1L in R? Show an example of creating each data type.
```{r problem_2, error=TRUE}
#Your code here
#The difference between 1 and 1L in R is that 1L is an integer while 1 is numeric(double).

print(typeof(1))
print(typeof(1L))
```

### Problem 3
Convert a numeric vector x <- c(1, 2, 3, 4) into a character vector. Why does x not change without assigning the coercion into a new object?
```{r problem_3, error=TRUE}
#Your code here
x <- c(1, 2, 3, 4)
x1 <- as.character(x)
print(x)
print(x1)
#x does not change without assigning the coercion into a new object because x is already in being used for a vector.
```

### Problem 4
Create a 3x4 matrix with numbers from 1 to 12. Assign row names ("Row1", "Row2", "Row3") and column names ("Col1", "Col2", "Col3", "Col4") to it.
```{r problem_4, error=TRUE}
#Your code here
rnames <- c("Row1", "Row2", "Row3")
cnames <- c("Col1", "Col2", "Col3", "Col4")
matrixA <- matrix(1:12, nrow = 3, ncol = 4, dimnames = list(rnames, cnames))
print(matrixA)
```

### Problem 5
Create a factor pet (pet is the object) with three levels: "Dog", "Cat", and "Bird".
```{r problem_5, error=TRUE}
#Your code here
pet <- as.factor(c("Dog", "Cat", "Bird"))
print(pet)
print(class(pet))
```

### Problem 6
How can you test if a vector v <- c(1, 2, NA, 3, 4, NaN, NaN, 4, NA) has any missing (NA) values or NaN values? Show how to find which elements are Na and NaN. What is the difference between NA and NaN?
```{r problem_6, error=TRUE}
#Your code here
v <- c(1, 2, NA, 3, 4, NaN, NaN, 4, NA)
is.na(v)
#NA represenets "not available" while NaN represents "not a number".
```

### Problem 7
Create a data frame named df with two columns: Name (with values "Alice", "Bob", "Charlie") and Age (with values 24, 30, 35).
```{r problem_7, error=TRUE}
#Your code here
df <- data.frame(ID = 1:3, Name = c("Alice", "Bob", "Charlie"), Age = c(24, 30, 35))
print(df)

```

### Problem 8
Assign names "A", "B", "C" to a numeric vector y <- c(2.5, 3.6, 1.8). Note: use the names() function.
```{r problem_8, error=TRUE}
#Your code here
y <- c(2.5, 3.6, 1.8)
names(y) <- c("A", "B", "C")
print(y)
```

### Problem 9
What will be the result of this operation: y <- c(1.7, "a")? Why?
```{r problem_9, error=TRUE}
#Your code here
y <- c(1.7, "a")
print(y)
print(typeof(y))
#When printing y, it results in charcters 
```

### Problem 10
You have a character vector z <- c("a", "b", "c"). What will be the result of as.numeric(z)?
```{r problem_10, error=TRUE}
#Your code here
z <- c("a", "b", "c")
as.numeric(z)
#It prints NA because you can't make letters numeric
```

### Problem 11
What does the vector() function do in R? Create an empty numeric vector of length 5 using vector().
```{r problem_11, error=TRUE}
#Your code here
#The vector function allows you to create a vector with a specific type and length
my_vector <- vector(mode = "numeric", length = 5)
print(my_vector)
```

### Problem 12
Create a list named my_list containing a numeric vector, a character vector and a logical value. The numeric vector should be the numbers 1-5, the character vector should be "A", "B", "C", and the logical vector is a single TRUE value.
```{r problem_12, error=TRUE}
#Your code here
my_list <- list(c(1:5), c("A", "B", "C"), TRUE)
```

### Problem 13
What is coercion in R? What will the vector v <- c(TRUE, 2) return?
```{r problem_13, error=TRUE}
#Your code here
#Coercion is the deliberate conversion of an object from one data type to another.
v <- c(TRUE, 2)
typeof(v)
```

### Problem 14
Concatenate two vectors x <- c(1, 2, 3) and y <- c(4, 5, 6) into a new vector. Combine x and y by row to create a matrix. Combine x and y by column to create a matrix
```{r problem_14, error=TRUE}
#Your code here
x <- c(1, 2, 3)
y <- c(4, 5, 6)
row_bind <- rbind(x, y)
col_bind <- cbind(x, y)
print(row_bind)
print(col_bind)

```

### Problem 15
You have a character vector text <- c("one", "two", "three"). Use the as.numeric() function on this vector and explain the output.
What is the hierarchy of coercions for logical, characters, numeric, integers, factors, and complex numbers?
```{r problem_15, error=TRUE}
#Your code here
text <- c("one", "two", "three")
as.numeric(text)
#Although the text represents numbers, it is written in text for instead of numeric so it is not able to be transformed from character to numeric like "3
```

### Problem 16
Use the mtcars dataset to graph a bar plot and histogram grouped by cyl (cylinders) variable.
```{r problem_16, error=TRUE}

# You can directly access mtcars data set in R
mtcars1 = mtcars
#Your code here
barplot(mtcars$cyl, main = "Bar Chart of Cylinders", xlab = "Cylinders", ylab = "Frequency")
```

### Problem 17
Using the movies.csv dataset (Canvas -> Files -> Homework -> Homework 2), create a scatter plot with the year for the y-axis and total gross revenue for the x-axis.
```{r problem_17, error=TRUE}
#Your code here
plot(movies$`Worldwide Gross`, movies$Year, main = "Gross Revenue per Year", xlab = "Gross Revenue", ylab = "Year")

```

### Problem 18
Install and load ggplot2 and create a plot for movies.csv dataset using this. What you graph is up to you, but use some of the special features associated with the package.
```{r problem_18, error=TRUE}
#Your code here
install.packages("ggplot2")
library(ggplot2)
ggplot(movies, aes(x = Year, y = Profitability)) + geom_point(color = "turquoise", size = 2) + ggtitle("Profitability per Year")
```